---
title: "p8105_hw1_rl3401.Rmd"
author: "Ruoxi Li"
date: "2023-09-22"
output: github_document
---
```{r,echo=FALSE}
Sys.setenv(LANGUAGE = "en")
```
## Problem 1
```{r,message=FALSE}
#set up the packages and the dataset
library(moderndive)
library(tidyverse)
data("early_january_weather")
```
1. the variables in the dataset `early_january_weather` are:

   `r names(early_january_weather)`.
   Besides, from the help document, we can see that

2. This dataset has **`r nrow(early_january_weather)` rows** and **`r ncol(early_january_weather)` columns**.


3. The mean temperature is **`r mean(pull(early_january_weather,temp))` F**.

```{r,message=FALSE}
#create the scatterplot
ggplot(early_january_weather, aes(x = time_hour, y= temp, color=humid)) + 
geom_point()
#save the scatterplot
ggsave("scatter1.png")
```
This scatterplot shows the relationship between `time_hour` on the **x-axis** and `temperature` on the **y-axis**, with the **color** of the points  based on the `humidity`. 

From the plot, it appears that as `time_hour` increases, there is a general tendency for both `temperature` and `humidity` to increase.

## Problem 2

```{r}
#create the dataframe
hw1_df <- tibble(
  random_sample = rnorm(10),
  logical_vector = random_sample > 0, 
  character_vector = c("R", "L", "3", "4", "0", "1", "_", "H", "W", "1"),  
  factor_vector = factor(rep(c("Level 1", "Level 2", "Level 3"),length = 10), levels = c("Level 1", "Level 2", "Level 3"))
)
```
```{r}
#show the dataframe
print(hw1_df)
```

```{r}
#calculate the mean of each variable
mean(pull(hw1_df,random_sample))
mean(pull(hw1_df,logical_vector))
mean(pull(hw1_df,character_vector))
mean(pull(hw1_df,factor_vector))
```

As shown above, the means of the `random sample` and the `logical factor` **can be calculated**, while the means of the `character vector` and the `factor vector` **can not be calculated**.

```{r,eval=FALSE}
mean(as.numeric(pull(hw1_df,logical_vector)))
mean(as.numeric(pull(hw1_df,character_vector)))
mean(as.numeric(pull(hw1_df,factor_vector)))
```

1. For the `logical vector`, the logical value TRUE is represented as 1, while FALSE is represented as 0.
 
   When the `as.numeric() function` is used, it returns the numerical values of 0 and 1 corresponding to the logical values.

2. It is not possible to calculate the mean of a character variable in R.

   When the `as.numeric() function` is applied to the `character vector`, R is unable to convert the character data type to a numeric data type, resulting in NA.

3. Finding the mean of a factor variable is also impossible. 

   When the `as.numeric() function` is used on the `factor vector`, R assigns numerical values of 1, 2, 3 to each level of the factor variable and the mean can be calculated.